Paper Analysis

### 2025-01-24 Paper_WriteUp/Paper_Plots.qmd
add some supplement plots

all_data_path <- rep(NA, 4)
all_data_path[1] <- here("box_data/lichess1700-1900/")
all_data_path[2] <- here("box_data/lichess2000-2200/")
all_data_path[3] <- here("box_data/lichess2300-2500/")
all_data_path[4] <- here("box_data/lichessGrandmasters/")
res_data_path <- rep(NA, 4)
res_data_path[1] <- here("results/lichess1700-1900/")
res_data_path[2] <- here("results/lichess2000-2200/")
res_data_path[3] <- here("results/lichess2300-2500/")
res_data_path[4] <- here("results/lichessGrandmasters/")

@@

```{r figure 3 left - plot global parameters for 17-19}
data_path <- here("results/lichess1700-1900")

fit_17_19_bullet_prev <- readRDS(file = 
                                   paste0(data_path,
                                          "/all_rated_bullet_model_prev.RDS"))

@@

```{r figure 4 left}
winner_plot_17_19 <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated W/L-Effect", y = element_blank(), 
       title = "1700-1900 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

@@

```{r plot player effects}
#| eval: false
indiv_plot_17_19 <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated Player Effect", y = element_blank()) +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

ggsave(plot = indiv_plot_17_19, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "17_19_prev_indiv.png")), dpi = 600,
       height = 7, width = 7)
```

@@

```{r figure 3 right - global parameters for gm data}
data_path <- here("results/lichessGrandmasters")

fit_gm_bullet_prev <- 
  readRDS(file = paste0(data_path,"/all_rated_bullet_model_prev.RDS"))

@@

## then the winner effects plot
winner_plot_gm <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated W/L-Effect", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

@@

```{r plot player effects}
#| eval: false
## then repeat for player effects
indiv_plot_gm <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated Player Effect", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

@@

ggsave(plot = indiv_plot_gm, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_prev_indiv.png")), dpi = 600,
       height = 7, width = 7)
```

@@

```{r figure 5 left - compare global params n-1 vs n-10}
## global effects first
data_path <- here("results/lichessGrandmasters")
fit_gm_bullet_10 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model.RDS"))

@@

hist_comp_winner <- player_pars_pre |> 
  filter(param == 2) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_pre_10 |> filter(param == 2) |> 
              mutate(y = as.factor(player_name))) |> 
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "D")

@@

## then player effects
## should update this also
hist_comp_player <- player_pars_pre |> 
  filter(param == 1) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_pre_10 |> filter(param == 1) |> 
              mutate(y = as.factor(player_name))) |> 
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "D")

@@

## Potential Supplement Plots
data_path <- here("results/lichess1700-1900")
fit_17_19_bullet_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_bullet_model_prev.RDS"
))
bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
fit_17_19_blitz_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_blitz_model_prev.RDS"
))
blitz_17_19 <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))

@@

comp_17_19_winner <- player_pars_bullet |> 
  filter(param == 2) |> 
  filter(player_name %in% common_players) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_blitz |>
              filter(param == 2) |>
              filter(player_name %in% common_players) |>
              mutate(y = as.factor(player_name))) |>
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "D")

@@

comp_17_19_player <- player_pars_bullet |> 
  filter(param == 1) |> 
  filter(player_name %in% common_players) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_blitz |>
              filter(param == 1) |>
              filter(player_name %in% common_players) |>
              mutate(y = as.factor(player_name))) |>
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "D")

@@

```{r figure 5 right - bullet v blitz gm}
data_path <- here("results/lichessGrandmasters")
fit_gm_bullet_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_bullet_model_prev.RDS"
))
bullet_gm <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
fit_gm_blitz_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_blitz_model_prev.RDS"
))
blitz_gm <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))
## compare global effects first
bullet_pars <- fit_gm_bullet_prev$draws(c(
  "mu_beta",
  "gamma1", "gamma2"
)) |>
  as_draws_df() |>
  pivot_longer(cols = "mu_beta":"gamma2") |>
  mutate(model = "Bullet")

@@

<!-- TO DO - Update Below here -->
<!-- Everything down as far as here up to date. -->
<!-- setting everything below to eval: false for now -->
<!-- Updating everything below here -->

@@

@@

## 2000-2200 Previous Game

```{r}
#| eval: false
data_path <- here("results/lichess2000-2200")

fit_20_22_bullet_prev <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))

fit_20_22_bullet_prev$summary(variables = c("mu_beta",
                                            "gamma1", "gamma2",
                                            "sigma_1", "tau[1]", "tau[2]",
                                            "sigma_g1", "sigma_g2"))
```

## 2300-2500 Previous Game

```{r}
#| eval: false
data_path <- here("results/lichess2300-2500")

fit_23_25_bullet_prev <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))

fit_23_25_bullet_prev$summary(variables = c("mu_beta",
                                            "gamma1", "gamma2",
                                            "sigma_1", "tau[1]", "tau[2]",
                                            "sigma_g1", "sigma_g2"))
```

@@

## Model Checking for the Other Cohorts

<!-- What about for the other two groups? -->

<!-- __To be added in__ -->

<!-- ![Time-Varying PPC for 2000-2200](../results/lichess2000-2200/time/ppc_fit_first_last_bullet.png) -->

<!-- ![Time-Varying PPC for 2300-2500](../results/lichess2300-2500/time/ppc_fit_first_last_bullet.png) -->

<!-- Maybe some very limited evidence that this becomes less -->
<!-- reasonable for players with a higher rating. -->

>>>>

all_data_path <- rep(NA, 4)
all_data_path[1] <- here("box_data/lichess1700-1900/")
all_data_path[2] <- here("box_data/lichess2000-2200/")
all_data_path[3] <- here("box_data/lichess2300-2500/")
all_data_path[4] <- here("box_data/lichessGrandmasters/")
res_data_path <- rep(NA, 4)
res_data_path[1] <- here("results_2025/lichess1700-1900/")
res_data_path[2] <- here("results_2025/lichess2000-2200/")
res_data_path[3] <- here("results_2025/lichess2300-2500/")
res_data_path[4] <- here("results_2025/lichessGrandmasters/")

@@

```{r figure 3 left - plot global parameters for 17-19}
data_path <- here("results_2025/lichess1700-1900")

fit_17_19_bullet_prev <- readRDS(file = 
                                   paste0(data_path,
                                          "/all_rated_bullet_model_prev_n1.RDS"))

@@

```{r figure 4 left}
winner_plot_17_19 <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |> 
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Change in Win Probability", y = element_blank(), 
       title = "1700-1900 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

@@

```{r figure s4 left - plot player effects}
#| eval: false
indiv_plot_17_19 <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated Player Effect", y = element_blank()) +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

ggsave(plot = indiv_plot_17_19, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S4_left.png")), dpi = 600,
       height = 7, width = 7)
```

@@

```{r figure 3 right - global parameters for gm data}
data_path <- here("results_2025/lichessGrandmasters")

fit_gm_bullet_prev <- 
  readRDS(file = paste0(data_path,"/all_rated_bullet_model_prev_n1.RDS"))

@@

## then the winner effects plot
winner_plot_gm <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |>
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Estimated W/L-Effect", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

@@

```{r figure s4 right - plot player effects}
#| eval: false
## then repeat for player effects
indiv_plot_gm <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated Player Effect", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

@@

ggsave(plot = indiv_plot_gm, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S4_right.png")), dpi = 600,
       height = 7, width = 7)
```

@@

```{r figure 5 left - compare global params n-1 vs n-10}
## global effects first
data_path <- here("results_2025/lichessGrandmasters")
fit_gm_bullet_10 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))

@@

hist_comp_winner <- player_pars_pre |> 
  filter(param == 2) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_pre_10 |> filter(param == 2) |> 
              mutate(y = as.factor(player_name))) |> 
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

@@

## then player effects
## should update this also
hist_comp_player <- player_pars_pre |> 
  filter(param == 1) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_pre_10 |> filter(param == 1) |> 
              mutate(y = as.factor(player_name))) |> 
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

@@

## Potential Supplement Plots
data_path <- here("results_2025/lichess1700-1900")
fit_17_19_bullet_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_bullet_model_prev_n1.RDS"
))
bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
fit_17_19_blitz_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_blitz_model_prev_n1.RDS"
))
blitz_17_19 <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))

@@

comp_17_19_winner <- player_pars_bullet |> 
  filter(param == 2) |> 
  filter(player_name %in% common_players) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_blitz |>
              filter(param == 2) |>
              filter(player_name %in% common_players) |>
              mutate(y = as.factor(player_name))) |>
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

@@

comp_17_19_player <- player_pars_bullet |> 
  filter(param == 1) |> 
  filter(player_name %in% common_players) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_blitz |>
              filter(param == 1) |>
              filter(player_name %in% common_players) |>
              mutate(y = as.factor(player_name))) |>
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

@@

```{r figure 5 right - bullet v blitz gm}
data_path <- here("results_2025/lichessGrandmasters")
fit_gm_bullet_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_bullet_model_prev_n1.RDS"
))
bullet_gm <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
fit_gm_blitz_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_blitz_model_prev_n1.RDS"
))
blitz_gm <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))
## compare global effects first
bullet_pars <- fit_gm_bullet_prev$draws(c(
  "mu_beta",
  "gamma1", "gamma2"
)) |>
  as_draws_df() |>
  pivot_longer(cols = "mu_beta":"gamma2") |>
  mutate(model = "Bullet")

@@

@@

<!-- TO DO - Tidy up below here -->

@@

## 2000-2200 Previous Game Winner/Loser Effect

```{r figure s1-left}
#| eval: false
data_path <- here("results_2025/lichess2000-2200")

fit_20_22_bullet_prev <- readRDS(file = 
                                   paste0(data_path,
                                          "/all_rated_bullet_model_prev_n1.RDS"))

bullet_20_22 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

player_id <- as_tibble(bullet_20_22) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- fit_20_22_bullet_prev$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id")

winner_plot_20_22 <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |> 
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Change in Win Probability", y = element_blank(), 
       title = "2000-2200 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

ggsave(plot = winner_plot_20_22, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S1_left.png")), dpi = 600,
       height = 7, width = 7)
```

## 2300-2500 Previous Game Winner/Loser Effect

```{r figure-s1-right}
#| eval: false
data_path <- here("results_2025/lichess2300-2500")

fit_23_25_bullet_prev <- readRDS(file = 
                                   paste0(data_path,
                                          "/all_rated_bullet_model_prev_n1.RDS"))

bullet_23_25 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

player_id <- as_tibble(bullet_23_25) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- fit_23_25_bullet_prev$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id")

winner_plot_23_25 <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |> 
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Change in Win Probability", y = element_blank(), 
       title = "2300-2500 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

ggsave(plot = winner_plot_23_25, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S1_right.png")), dpi = 600,
       height = 7, width = 7)
```

@@

To be added here for 1700-1900 and GM cohorts.

```{r figure S3 left}
## global effects first
data_path <- here("results_2025/lichess1700-1900")

fit_17_19_bullet_5 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n5.RDS"))

fit_17_19_bullet_10 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))

prev_pars <- fit_17_19_bullet_prev$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. Game")

pre_5_pars <- fit_17_19_bullet_5$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. 5 Games")

pre_10_pars <- fit_17_19_bullet_10$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. 10 Games")

hist_comp_global <- bind_rows(prev_pars, pre_5_pars, pre_10_pars) |> 
  ggplot(aes(x = value, color = model)) +
  stat_pointinterval(position = position_dodge(width = 0.2)) +
  facet_wrap(~name,
             scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_3col() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

ggsave(plot = hist_comp_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S3_left.png")), dpi = 600,
       height = 3.5, width = 7)

```{r figure S3 right}
## global effects first
data_path <- here("results_2025/lichessGrandmasters")

fit_gm_bullet_5 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n5.RDS"))

fit_gm_bullet_10 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))

pre_5_pars <- fit_gm_bullet_5$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. 5 Games")

hist_comp_global <- bind_rows(prev_pars, pre_5_pars, pre_10_pars) |> 
  ggplot(aes(x = value, color = model)) +
  stat_pointinterval(position = position_dodge(width = 0.2)) +
  facet_wrap(~name,
             scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_3col() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

ggsave(plot = hist_comp_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S3_right.png")), dpi = 600,
       height = 3.5, width = 7)
